---
title: "R Notebook"
output:
  html_document:
    df_print: paged
---
### Load packages 
```{r include=FALSE}
library(ggplot2)
library(car)
library(lmtest)
library(plyr)
library(dplyr)

```

### Open Files
```{r include=FALSE}
Muzz = read.csv('./Lythrum/Eugene_Data')
Eugene = read.csv('./Lythrum/Eugene_Data')
PL_match_0 = read.csv("./Lythrum/PL_match_0")
PL_match_c = read.csv("./Lythrum/PL_match_c")
PL_miss_0 = read.csv("./Lythrum/PL_miss_0")
PL_miss_c = read.csv("./Lythrum/PL_miss_c")
```

### Check to see if data is normal
```{r}
mod1 = lm(Damage~LAT+VEGWT+RWT+Stem.Length+Stem.Count+RWT, data=Eugene)
car::qqp(Eugene$Damage, "norm")
#plot(density(Eugene$Damage)) # does it fit a normal distribution?
bptest(mod1) #check for homoskedastity
summary(mod1) 
#plot(mod1)
mod1.mVEGWT<-update(mod1,.~. - VEGWT)
anova(mod1,mod1.mVEGWT)

mod1.mLAT<-update(mod1.mVEGWT,.~. - LAT)
anova(mod1.mVEGWT,mod1.mLAT)

mod1<-mod1.mLAT
summary(mod1)
```

1. Stem.length is significantly correlated to Damage
2. Stem.Count is significantly correlated to Damage 
3. Previous years vegetative weight is significantly correlated to damage when looked at alone
4. Previous years root weight is slightly significantly correlated to Damage
5. No apparent correlation between LAT and Damage FOR NOW --- CORRELATION IS SHOWN LATER



### Group the values
```{r}
diff = max(Eugene$LAT, na.rm = T)-min(Eugene$LAT, na.rm = T)
cutoff = diff/3.0
Min_LAT =min(Eugene$LAT, na.rm = T)

# NA is evaluated as a value within the South range, so we remove those columns/ make them zeros
Eugene = Eugene[Eugene$FAMFULL!='P137',]
Eugene = Eugene[Eugene$FAMFULL!='P176',]

#Make new column Origin, group by LAT and replace with 1_North, 2_Mid and 3_South
Eugene$Origin = Eugene$LAT
Eugene$Origin[Eugene$Origin >= Min_LAT & Eugene$Origin<= Min_LAT+cutoff ] = '3_South'
Eugene$Origin[Eugene$Origin > Min_LAT+cutoff & Eugene$Origin <=Min_LAT+2*cutoff]= '2_Mid'
Eugene$Origin[Eugene$Origin > Min_LAT+2*cutoff]= '1_North'

```

### Group by FAMFULL, POP and Origin. 
###### Get means for groups 
```{r}
FAM_mean = ddply(Eugene,~FAMFULL,summarise,LAT_mean=mean(LAT),Dam_mean=mean(Damage), SL_mean=mean(Stem.Length), SC_mean = mean(Stem.Count), VEGWT_mean=mean(VEGWT,na.rm = T), RWT_mean=mean(RWT, na.rm=T), PREV_mean=mean(PREV, na.rm = T))

POP_mean = ddply(Eugene,~POP,summarise,LAT_mean=mean(LAT),Dam_mean=mean(Damage), SL_mean=mean(Stem.Length), SC_mean = mean(Stem.Count), VEGWT_mean=mean(VEGWT,na.rm = T), RWT_mean=mean(RWT, na.rm=T), PREV_mean=mean(PREV, na.rm = T))
write.csv(POP_mean,'./Lythrum/POP_mean.csv', row.names = F)

Origin_mean = ddply(Eugene,~Origin,summarise,LAT_mean=mean(LAT),Dam_mean=mean(Damage), SL_mean=mean(Stem.Length), SC_mean = mean(Stem.Count), VEGWT_mean=mean(VEGWT,na.rm = T), RWT_mean=mean(RWT, na.rm=T), PREV_mean=mean(PREV, na.rm = T))
```


### Analysis of ungrouped data
##### Damage VS...
```{r include=FALSE}
mod2 = lm(Damage~LAT*Stem.Length, data=Eugene)
summary(mod2)

mod3 = lm(log(Damage+1)~LAT*Stem.Length, data=Eugene)
summary(mod3)

mod4 = lm(Damage~LAT, data=PL_match_0)
summary(mod4)
```

##### Plant growth VS...
```{r}
mod7 = lm(Stem.Length~LAT, data=Eugene)
summary(mod7)

#TEST TO SEE IF PREVIOUS YEAR AFFECTED THIS YEAR GROWTH
mod_growthl = lm(Stem.Length~RWT*VEGWT, data=PL_match_0)
summary(mod_growthl)
mod_growthc = lm(Stem.Count~RWT*VEGWT, data=PL_match_0)
summary(mod_growthc)

# The previous years plant size/ growth is significantly correlated to this years stem lentgh and even more to stem count
```

### Grouped mean analysis
#### Damage VS...
```{r}
mod5 = lm(Dam_mean~VEGWT_mean, data=POP_mean)
summary(mod5) 
qplot(x=VEGWT_mean, y=Dam_mean, data=POP_mean)

mod6 = lm(Dam_mean~SL_mean, data=POP_mean)
summary(mod6)
qplot(x=SL_mean, y=Dam_mean, data=POP_mean)

mod8 = lm(Dam_mean~VEGWT_mean*RWT_mean, data=POP_mean)
summary(mod8)

mod9 = lm(Dam_mean~LAT_mean, data=POP_mean)
summary(mod9)
qplot(x=LAT_mean, y=Dam_mean, data=POP_mean, colour=POP) +theme_classic()
# Has a peak LAT 



```
#### There is significant correlation between Damage and Stem Length when Grouped by Population
#### There is slightly significant between Damage and Vegetative weight from previous year when Grouped by Population


### PLOT CITY

Plot of Latitude mean against Damage mean for POP_mean

```{r}
qplot(x=SL_mean, y=Dam_mean, data=POP_mean, colour=POP) +theme_classic()

```

####Nice graph for Stem Length and Damage correlation


```{r}
qplot(x=VEGWT_mean, y=Dam_mean, data=POP_mean, colour=POP) +theme_classic()
```

Plot for Vegetative weight against Damage

```{r include=FALSE}
qplot(x=LAT, y=Stem.Length, data=PL_match_0) +theme_classic()
#Plot of Stem Length against LAT for PL_match
```

```{r include=FALSE}
qplot(x=LAT, y=Stem.Length, data=Eugene, colour=POP) +theme_classic()
#Plot of Stem Length against LAT for Eugene

```


```{r}
qplot(x=LAT_mean, y=SL_mean, data=POP_mean, colour=POP) +theme_classic()
```

Plot of Stem Length against LAT for POP_mean


```{r}
qplot(x=LAT_mean, y=Dam_mean, data=Origin_mean, colour = Origin, shape= Origin) +theme_classic()


```

Small but significant difference in Damage associated with the Mid Origin of the plants, I will look into the areas given by the Latitude groupings. I want to see if this latitude is similar to that of the natural/original habitat of the Lythrum, before it was introduced to north america.


##Mixed effects analysis
```{r Mixed effects}
library(lme4)
LMM1<-lmer(Stem.Length~LAT*Origin+(1|FAMFULL), data=Eugene)
summary(LMM1)
# Compare to a simple model
LMM0<-lmer(Stem.Length~1+(1|FAMFULL),data=Eugene)
summary(LMM0)
# Different among-family variances for each Origin
LMM2<-lmer(Stem.Length~Site*Origin+(Origin|FAMFULL),data=LyDat)
summary(LMM2)

# Do origins differ in among-family variance?
# Use a likelihood ratio test
anova(LMM2,LMM1)

```


## NMDS
```{r}

```


































